|

### 2020-12-19 manuscript/manuscript_1.0.Rmd
First draft of the methods done. Add a parameter, sharecost, determining which proportion of the cost is beared by the winner.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

We model the evolution of signals of quality in the context of agonistic interactions by 
modelling a population of haploid individuals with non-overlaping generations. 
Individuals are born every generation with a level of quality ($Q_i$, where $i$ is subscript 
of the population vector of size $N$) given by a number between zero and one, which is drawn 
from a truncated normal distribution $N(\mu = 0.5,\sigma)$. 
An Individual with quality $0$ has the lowest RHP, 
while an individual with quality $1$ has the highest. As part of development processes
individuals produce a morphological signal (badge), the size of such depends on their level 
of quality according to a reaction norm given by $B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$;
where are $\alpha_i$ and $\beta_i$ are individual specific traits that determine 
the shape of the reaction norm. Variation in $\alpha_i$ and $beta_i$ means individuals
can have either uninformative (flat) or informative (logistic) norms; but the size of 
the signal is constrained between zero and one. We assume different values of these traits 
are given by different alleles, and are inherited from mother to offspring. 
Unless, with a small probability, mutation changes the allelic value of the offspring.
After birth individuals go through a round of viavility selection. 
Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)}$. 
Where $k_1$ and $k_2$ are parameter values determining the shape of the survival function. 
Importantly, if $k_2=0$ survival probability is independent of quality, while if $k_2>0$
lower quality individuals pay a higher price for s similar-sized badge, fulfilling the core assumption 
of the handicap principle [@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].

@@

The decision of whether to escalate a fight against an interacting partner can be dependent 
on the size of the partner´s badge, the actual dependency is determined by 
the focal´s experiences through a learning process. We implement learning using 
the actor-critic aproach from reinforcement learning theory [RL; @sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
Individuals estimate the reward (payoff) expected from  interacting with partners 
of different badge sizes (the critic in RL terminology). 
After every interaction they update the estimate of reward proportionally to the difference between their current estimate and the observed reward (prediction error $\delta$). Furthermore, they express different probabilities of escalatining (retreating) depending on
the badge size of their opponent (the actor in RL). They update the probability of 
retreating (escalating)  up or down depending on whether retreating (escalating) leads to an increase 
in the reward estimation. So, if a focal individual decides to escalate a fight against 
an individual with small badge, and that leads to an increase in the reward estimation,
the focal individual will increase the probability of escalating fights with 
individuals of small badges in the future. Given that badge size is a real number between 0 and 1,
the reward estimation, as well as the probability of retreating (escalating), must be
generalized across different values. To implement generalization we use the function
approximation method based on radial basis functions [@sutton_Reinforcement_2018]. Specifically, we pick $c$ evenly spaced 
values along the badge size interval,

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(png)
require(here)
source(here("../R_files/posPlots.r"))
source(here("AccFunc.r"))
```

@@

We model the evolution of signals of quality in the context of agonistic interactions by 
modelling a population of haploid individuals with non-overlaping generations. 
Individuals are born every generation with a level of quality ($Q_i$, where $i$ is subscript 
of the population vector of size $N$) given by a number between zero and one, which is drawn 
from a truncated normal distribution $N(\mu = 0.5,\sigma)$. 
An individual with quality $0$ has the lowest RHP, 
while an individual with quality $1$ has the highest. As part of development processes
individuals produce a morphological signal (badge), the size of such depends on their level 
of quality according to a reaction norm given by $B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$;
where are $\alpha_i$ and $\beta_i$ are individual specific traits that determine 
the shape of the reaction norm. Variation in $\alpha_i$ and $\beta_i$ means individuals
can have either uninformative (flat) or informative (logistic) norms; but the size of 
the signal is constrained between zero and one. We assume different values of these traits 
are given by different alleles, and are inherited from mother to offspring. 
Unless, with a small probability, mutation changes the allelic value of the offspring.
After birth individuals go through a round of viavility selection. 
Individual specific survival probability is given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)}$. 
Where $k_1$ and $k_2$ are parameter values determining the shape of the survival function. 
Importantly, if $k_2=0$ survival probability is independent of quality, while if $k_2>0$
lower quality individuals pay a higher price for s similar-sized badge, fulfilling the core assumption 
of the handicap principle [@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].

@@

The decision of whether to escalate a fight against an interacting partner can be dependent 
on the size of the partner´s badge, the actual dependency is determined by 
the focal´s experiences through a learning process. We implement learning using 
the actor-critic aproach from reinforcement learning theory [RL; @sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
Individuals estimate the reward (payoff) expected from  interacting with partners 
of different badge sizes (the critic in RL terminology). 
After every interaction they update the estimate of reward proportionally to the difference between their current estimate and the observed reward (prediction error $\delta$). Furthermore, they express different probabilities of escalatining (retreating) depending on
the badge size of their opponent (the actor in RL). They update the probability of 
retreating (escalating)  up or down depending on whether retreating (escalating) leads to an increase 
in the reward estimation. So, if a focal individual decides to escalate a fight against 
an individual with small badge, and that leads to an increase in the reward estimation,
the focal individual will increase the probability of escalating fights with 
individuals of small badges in the future. Given that badge size is a real number between 0 and 1,
the reward estimation, as well as the probability of retreating (escalating), must be
generalized across different values. To implement generalization we use the linear function
approximation method based on radial basis functions [@sutton_Reinforcement_2018]. Specifically, we pick $f$ feature centers, which are evenly spaced values along the badge size interval ([0,1]). Each one of these features centers is associated with a weight for reward estimation and 
tendency to play retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights associated 
with each feature, weighted by the response triggered by the feature. 
The response of each feature center dimishes as a gaussian function  with the distance between the feature center and the badge size of the partner. Formally the reward estimation 
$\hat{R}$ when the focal individual ($i$) faces individual $j$ is given by,

\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
\end{equation}

where $x_z$ is the weight of feature center $z$ on the reward estimation; and $\theta$ is
the width of the generalization function. Similarly, the tendency to retreat is given by sum 
of feature weigths, and the probability of associated with the tendency is obtained
by applying a logistic transformation. So formally, the log-odds to retreat
when facing individual $j$ is

\begin{equation}
logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-c_z|}{2\theta^2})} 
\end{equation}

where $y_z$ is the weight of feature center $z$ on the tendency to retreat.

After all their interactions, individuals in the population reproduce proportionally to 
their fitness $w_i$, which is a sum of the baseline fitness ($w_0$) and all the payoffs
obtained throughout their life. Thus, the combination of natural selection and genetic drift
changes the distribution of values in $\alpha$ and $\beta$ that segregates in the population, effectively the badges expressed and changing the communication system.

```{r Figure 1,echo=FALSE,fig.width=8,fig.align='center',fig.height=4,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual. Red shows an honest reaction norm, while the blue shows a dishonest reaction norm. B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward."}
honestBig<-readPNG(here("Images/cart_honestBig.png"))
honestSmall<-readPNG(here("Images/cart_honestSmall.png"))
dishonest<-readPNG(here("Images/cart_dishonest.png"))
par(plt=posPlot(numplotx = 2,idplotx = 1,upboundx = 91),las=1)
  rangPars<-data.frame(honest=c(4,2),dishonest=c(0,-1.8))
 rangQual<-seq(0,1,by = 0.01)
 linesSender<-sapply(rangPars,function(x)logist(rangQual,alpha = x[2],beta = x[1]))
 plot(0,0,type="l",xlab="",xlim=c(0,1),
      ylab ="",cex.lab=1,lwd=2,col="grey",cex.axis=1,ylim=c(0,1.1))
 text(x = 0.08,y=1.05,"A",cex = 1.5)
 mtext("Quality",1,line = 2,cex = 1,las = 0)
 matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
 mtext("Badge size",2,line = 2.3,cex = 1,las = 0)
 rasterImage(honestBig,0.75,0.6,0.95,0.8)
 rasterImage(honestSmall,0.05,0.01,0.25,0.21)
 rasterImage(dishonest,0.75,0.87,0.95,1.07)
 rasterImage(dishonest,0.05,0.87,0.25,1.07)
 # legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")

 par(plt=posPlot(numplotx = 2,idplotx = 2,upboundx = 91),las=1,new=TRUE)

 nCenters<-6
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
sigSq<-0.01
weights<-rep(0,nCenters)
weights<-c(-5,-2,-1,1,2,5)#runif(nCenters,min=-1,max=1)
rangx<-seq(0,1,length=1000)

plot(logist(totRBF(rangx,centers,sigSq,weights),alpha = 0,beta=1)~rangx,type='l',col=1,
     xlab="",ylab="",ylim=c(0,1.1),lwd=3,xaxt="s",cex.lab=2,cex.axis=1,yaxt='n')
axis(4,cex=1)
text(x = 0.08,y=1.05,"B",cex = 1.5)
# test<-sapply(1:6, function(x){RBF()})
points(y=logist(weights,alpha = 0,beta=1),x=centers,cex=1.5,pch=20)
respon<-logist(weights[4]*RBF(rangx,centers[4],0.01))
lines(x=rangx,y=respon,col="grey",lwd=2)
axis(side = 4,at = seq(range(respon)[1],range(respon)[2],length.out = 4),
     labels = round(seq(0,1,length.out = 4),2),line = -4,col="grey")

# xRBFs<-sapply(centers, function(x){seq(x-0.2,x+0.2,by=0.001)})
# RBFS<-sapply(1:6, function(x){
#   logist(weights[x]*RBF(seq(centers[x]-0.2,centers[x]+0.2,by=0.001),xCenter = centers[x],
#                         sigSq = 0.01))
# })
# matlines(x=xRBFs,y=RBFS,col="grey",lty=2,lwd=2)

mtext("prob. retreat",4,line = 2.3,cex = 1,las = 0)
mtext("Badge size",1,line = 2,cex = 1,las = 0)

```

## Results
